---
title: "Introduction to ggepi"
author: "Luke W. Johnston"
date: "`r Sys.Date()`"
output: 
    rmarkdown::html_vignette:
        toc: true
vignette: >
  %\VignetteIndexEntry{Introduction to ggepi}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.width = 7,
    fig.height = 7,
    fig.align = "center",
    warning = FALSE
    )
library(ggplot2)
library(ggepi)
```

## Estimates and confidence intervals from regression models

In most cases, showing a figure of some results is more impactful than showing a
table. Often times in epidemiology, results from analyses are put into a table.
This is less then ideal for busy researchers, as it takes longer to process and
understand the results from a table than it is from a figure. So

```{r}
library(broom)
fit <- lm(Fertility ~ 0 + Catholic + Agriculture + Examination + Education + Infant.Mortality, data = swiss)
fit <- tidy(fit, conf.int = TRUE)
```

```{r}
view_estci(fit)
```

If you want more fine-grained control, you can use the `geom_estci` layer directly. 

```{r}
ggplot(fit, aes(x = estimate, y = term, xmin = conf.low, xmax = conf.high)) +
    geom_estci()
```

There is a utility function `discrete_pvalue` to convert a numeric p-value into a factor, which makes it easier to visualize the results. So:

```{r}
fit2 <- transform(fit, p.value = discrete_pvalue(fit$p.value))

p <- ggplot(fit2, aes(x = estimate, y = term, xmin = conf.low, xmax = conf.high,
                 alpha = p.value, size = p.value)) 
p + geom_estci(fatten = 1, center.linetype = "dotted")
```

We can add ends to the error bars too.

```{r}
p + geom_estci(fatten = 1, center.linetype = "dotted", height = 0.5, linetype = "dashed")
```

All of the power of ggplot2 is available to customize the `geom_estci` layer.
Check out the documentation for the `geom_estci` function as well as the ggplot2
layers e.g. `geom_pointrange` or `geom_point` for full customization. Since this
is a ggplot2 object, you can add to it all you want, like using `facet_grids`.

# Similar packages

There are a few packages that have similar functionality as the e.g.
`geom_estci`, but each has its own specific use cases:

- [dotwhisker](https://cran.r-project.org/package=dotwhisker) package
- [GGally](https://ggobi.github.io/ggally/) package

In both of these packages, they assume you'll plot *directly* from the `lm` or
other regression function, so they do a lot of internal data wrangling and
managing. The ggepi package is as hands off as possible when it comes to your
data you plan to plot, so all of the work in getting the data ready for the 
figure must be done before hand. This package assumes what you want to plot is
what you want to plot, nothing more. There are trade offs to both ways of doing
things! Try the different packages out yourself and find out what works for you!
